Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection

### 2017-03-08 ccppabc/text/revision/main.tex
Further edits to Section 3.4 to address Ref Report #1

In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours enclosing the $68\%$ 
and $95\%$ confidence regions of the posterior probabilities of the two methods. Figure \ref{fig:cont_nbarxi}
uses $\ngalaxy$ and $\xigg$ as observables, while Figure \ref{fig:cont_nbargmf} uses $\ngalaxy$ and $\gmf$ 
as observables. 
In both figures, the ``true'' HOD parameters are plotted as black stars. The overall shape of the contours 
are generally in good agreement with each other. However, the contours for the ABC-PMC method are
more elongated along $\alpha$.

%%%
{\bf \color{dred} 
Overall, the constraints from ABC-PMC are consistent with those from 
the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
advantages. ABC-PMC utilizes a forward generative model. Our forward generative model, 
for instance, accounts for sample variance. Forward generative models also have the 
advantage that they can account for sources of systematic uncertainties that affect 
observational data.
}
%%%

%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
%MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
%ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
%accounts for sample variance. In principle, they also have the advantage that they can account for 
%sources of systematic uncertainties in the observational data.

%%%
{\bf \color{darkgreen} 
CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
matrix depends on the accuracy of the mock catalogs (a.k.a. the forward model). While it's definitely 
true that on small scales they break down, forward models in ABC suffers from the same issues.
} 
%%%
Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix
estimate that captures the sample variance of the data. While we are able to do this accurately 
within the scope of the HOD framework, for more general LSS parameter inference situations, it 
is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs,
which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein). 
Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a 
covariance matrix estimate; hence it does not face these problems.

{\em ABC-PMC -- unlike the Gaussian-likelihood approach -- is agnostic about the functional form of the 
underlying distribution of the summary 
statistics} ($\xigg$ and $\gmf$). This may explain why we find less biased constraints from the
ABC-PMC analysis compared to the Gaussian-likelihood analysis, when using $\gmf(N)$ as an observable 
(Figures \ref{fig:hist_nbargmf} and \ref{fig:cont_nbargmf}).

>>>>

{\bf \color{dred}
In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours 
enclosing the $68\%$ and $95\%$ confidence regions of the posterior probabilities of 
the two methods using $[\ngalaxy, \xigg]$ and $[\ngalaxy, \gmf]$ as observables 
respectively. In both figures, we mark the ``true'' HOD parameters (black star). The 
overall shape of the contours are in agreement with each other. However, we note
that the contours for the ABC-PMC method are more extended along $\alpha$. 
}

%%% RefReport #1 
%%% CHH (3/8/2017): Made significant changes to the text in order to better emphasize
%%%      the advantages of ABC over MCMC
{\bf \color{dred} 
Overall, the HOD parameter constraints from ABC-PMC are consistent with those from 
the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
advantages. For instance, ABC-PMC utilizes a forward generative model. Our forward 
generative model accounts for sample variance. On the other hand, the Gaussian 
pseudo-likelihood approach, as mentioned earlier this section, does not account 
for sample variance in the model and relies on the covariance matrix estimate to 
capture the sample variance of the data. 
}

{\bf \color{dred} 
Accurate estimation of the covariance matrix 
in LSS, however, faces a number of challenges. It is both labor and computationally 
expensive and dependent on the accuracy of simulated mock catalogs, known to be 
unreliable on small scales~(see \citealt{cosmiccode,nifty} and references therein). 
In fact, as \cite{Sellentin:2016a} points out, using estimates of the covariance 
matrix in the Gaussian psuedo-likelihood approach become further problematic. Even 
when inferring parameters from a Gaussian-distributed data set, using covariance 
matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
function that is {\em no longer} Gaussian.
ABC-PMC does not depend on a covariance matrix estimate; hence, it does not face 
these problems. In principle, forward models of the ABC-PMC method also have the 
advantage that they can account for sources of systematic uncertainties that affect 
observations.
}
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
%Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix estimate that captures the sample variance of the data. While we are able to do this accurately within the scope of the HOD framework, for more general LSS parameter inference situations, it is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein).  Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a covariance matrix estimate; hence it does not face these problems.

Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
is agnostic about the functional form of the underlying distribution of the 
summary statistics} 
($\xigg$ and $\gmf$). 
{\bf \color{darkgreen}
CHH: Need a sentence here that reiterates that the Gaussian functional form 
assumption is incorrect and especially incorrect for the GMF observable case.
Also, for us to say the following sentence, we need to talk about it more explicitly 
when discussing Figures 7,8,9, and 10. 
}
This may explain why we find less biased constraints from the
ABC-PMC analysis compared to the Gaussian-likelihood analysis, when using $\gmf(N)$ as an observable 
(Figures \ref{fig:hist_nbargmf} and \ref{fig:cont_nbargmf}).
